\documentclass{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{geometry}
\usepackage{xcolor}

\geometry{margin=1.15in}

\title{Proposed Paper Update: SBO-Lite With Qualitative Regression Guard}
\author{}
\date{}

\begin{document}
\maketitle

\section{Motivation}

The current implementation of SBO-Lite uses the qualitative verifier as a
guardrail against obvious critique regressions. This differs from the latest
paper draft, where the verifier only reports labels
\(\{\mathrm{resolved}, \mathrm{partial}, \mathrm{unresolved}\}\) and does not
participate directly in accept/reject decisions. The implementation is arguably
more useful for small local models and short validation sets: a candidate may
improve the current validation batch while visibly undoing an important stored
critique. In that setting, a qualitative regression guard is a practical
stability mechanism.

This note proposes a paper-side update that makes SBO-Lite faithful to the
implemented behavior while keeping the main distinction from full SBO clear:
SBO-Lite does not use calibrated numerical judge scores, centered baselines
\(s_i^0\), loss-scale parameters \(\lambda_i\), or the semantic max-envelope
model \(M_k\).

\section{Revised SBO-Lite Verifier Contract}

For a candidate prompt \(p\), current center \(\hat p\), active critique bundle
\(\mathcal A_k\), and watchlist \(\mathcal W_k\), the verifier returns a
qualitative JSON object:
\[
    V_k(p)
    =
    \{\mathrm{active}, \mathrm{watchlist}, \mathrm{blocking\_regression},
    \mathrm{summary}\}.
\]
Each active or watchlist item receives one of three labels:
\[
    r_i(p)\in
    \{\mathrm{resolved},\mathrm{unclear},\mathrm{regressed}\}.
\]
Here:
\begin{itemize}
    \item \(\mathrm{resolved}\): the candidate clearly fixes or preserves the
    critique relative to the current prompt;
    \item \(\mathrm{unclear}\): the candidate is unrelated to the critique or
    only partially addresses it;
    \item \(\mathrm{regressed}\): the candidate reintroduces or worsens the
    failure mode described by the critique.
\end{itemize}

The field \(\mathrm{blocking\_regression}\) is true if the verifier identifies
at least one serious regression on the active bundle or watchlist. This field is
not a calibrated mathematical constraint; it is a practical guardrail.

\section{Revised Controller}

Let \(\widehat F_k(p)\) be the finite validation estimate used by the
controller, and define the observed improvement
\[
    \widehat\Delta_k(p)
    =
    \widehat F_k(\hat p)-\widehat F_k(p).
\]
SBO-Lite forms the admissible candidate set
\[
    \mathcal G_k^{\mathrm{lite}}
    =
    \{p\in\mathcal C_k:
      \widehat\Delta_k(p)>0
      \ \mathrm{and}\
      \mathrm{blocking\_regression}(p)=\mathrm{false}\}.
\]
If \(\mathcal G_k^{\mathrm{lite}}\) is nonempty, the controller chooses
\[
    p_k^{\star}\in
    \arg\min_{p\in\mathcal G_k^{\mathrm{lite}}}\widehat F_k(p).
\]
If the set is empty, the center remains fixed and the best evaluated candidate
is used only to refine the critique bundle:
\[
    p_k^{\mathrm{refine}}\in
    \arg\min_{p\in\mathcal C_k}\widehat F_k(p).
\]

\section{Algorithm Text Replacement}

\begin{algorithm}[H]
\caption{SBO-Lite With Qualitative Regression Guard}
\label{alg:sbo_lite_guard}
\begin{algorithmic}[1]
\STATE \textbf{Input:} initial prompt \(p_0\), proposer \(\pi\), evaluator,
qualitative verifier, candidate budget \(N\), stopping patience \(H\).
\STATE \textbf{Initialize:} center \(\hat p=p_0\), critique bundle
\(\mathcal B_0=\{c_0\}\), null counter \(h=0\).
\FOR{\(k=1,\ldots,T\)}
    \STATE Select active critiques \(\mathcal A_k\) and watchlist critiques
    \(\mathcal W_k\) from \(\mathcal B_k\).
    \STATE Generate candidates
    \(\mathcal C_k=\{p_k^{(1)},\ldots,p_k^{(N)}\}\) using
    \(\pi(\cdot\mid \hat p,\mathcal A_k,\mathcal W_k)\).
    \STATE For every \(p\in\mathcal C_k\), evaluate \(\widehat F_k(p)\) and
    ask the verifier for labels against \(\mathcal A_k\cup\mathcal W_k\).
    \STATE Form
    \[
        \mathcal G_k^{\mathrm{lite}}
        =
        \{p\in\mathcal C_k:
        \widehat F_k(p)<\widehat F_k(\hat p)
        \text{ and no blocking regression is detected}\}.
    \]
    \IF{\(\mathcal G_k^{\mathrm{lite}}\neq\emptyset\)}
        \STATE \textbf{Serious Step:} choose the candidate in
        \(\mathcal G_k^{\mathrm{lite}}\) with lowest \(\widehat F_k\), set it
        as the new center, reset \(h\leftarrow0\), and generate/update
        critiques from remaining validation failures.
    \ELSE
        \STATE \textbf{Null Step:} keep the center fixed, set
        \(h\leftarrow h+1\), and add or promote a critique explaining either
        failed validation improvement or a verifier-detected regression.
    \ENDIF
    \STATE Update \(\mathcal B_{k+1}\) by keeping active unresolved or
    regressed critiques, newly discovered failure modes, and a small watchlist
    of historically important critiques.
    \STATE Stop if \(h\ge H\), no candidate improves validation loss, or the
    evaluation budget is exhausted.
\ENDFOR
\end{algorithmic}
\end{algorithm}

\section{Suggested Paper Wording}

SBO-Lite should be described as a qualitative guarded controller rather than a
calibrated bundle model. The verifier is not estimating a numeric semantic
inner product and has no theorem-level role analogous to \(M_k\). Instead, it
provides a conservative regression flag. Acceptance remains validation-loss
first among candidates that pass this qualitative guard.

The full SBO theory should remain attached only to the calibrated algorithm:
centered semantic scores, positive loss-scale parameters, and the semantic
max-envelope model. SBO-Lite should be positioned as a practical agentic
variant whose stability is heuristic and traceable, not as a direct instance of
the finite-candidate semantic bundle validity theorem.

\end{document}
